---
title: "expct"
author: "Oisin Ryan"
date: "2022-12-15"
output: github_document
---

# expct

An `R` package to aid researchers in analyzing experience sampling data...

The uses XXX to do YYY, allowing researchers to:
  - a) 
  - b) quantify the uncertainty around those measures in the form of confidence/credible intervals 
  
## Background
This repository contains an `R` package used by ZZZ (in preperation0) to qqq

## Installation
The current version of this package can be installed directly from github using
```{r,  eval = F}
devtools::install_github("ryanoisin/CTVEM")
```

## Usage
Takes as input data in this format.. time is in this format. The main function is.... this outputs this...

### Example

```{r}
# load("data/simdata.rda")
library(CTVEM)
head(simdata)

```

```{r}
Tpred = seq(1,30,1)

est <- CTVEM(
  data = simdata,
  Time = "time",
  outcome = c("Y1","Y2"),
  ID = "id",
  estimate = "marginal",
  Tpred = Tpred,
  plot_show = F,
  quantiles = c(.025, 0.975),
  method = "bam",
)

plot(est$est[[1]], type = "b", ylab = "estimated autoregression", xlab = "time diff", main = "Autoregression X1")
lines(est$highCI[[1]], lty = 2, col = "gray")
lines(est$lowCI[[1]], lty = 2, col = "gray")

plot(est$est[[2]], type = "b", ylab = "estimated autoregression", xlab = "time diff", main = "Cross-Regression X2 on X1")
lines(est$highCI[[2]], lty = 2, col = "gray")
lines(est$lowCI[[2]], lty = 2, col = "gray")
# plot(mcr$est[[2]], type = "b", ylab = "estimated autoregression", xlab = "time diff", main = "Cross-Regresion")
# plot(mcr$est[[3]], type = "b", ylab = "estimated autoregression", xlab = "time diff", main = "Autoregression X1")
# plot(mcr$est[[4]], type = "b", ylab = "estimated autoregression", xlab = "time diff", main = "Autoregression X1")
```

Other output details... Other input details...


## Contact Details

For more details please contact **o.ryan@uu.nl**
