---
title: "expct"
author: "Oisin Ryan"
date: "2022-12-15"
output: 
  md_document:
    variant: gfm
---

# expct

An `R` package to estimate auto- and cross-correlations from time-series data sampled with any arbitrary sampling scheme. 

The package takes a time-series dataset with measurement timing information as the input. Relying on a simple data-stacking approach and Generalized Additive Mixed Models (GAMMs), it allows researchers to:

  - a) Estimate auto- and cross-correlations at different time-intervals requested by the user.
  - b) Visualize how lagged correlations vary and evolve as a function of the time-interval between measurements
  - c) Construct confidence intervals (CIs) to quantify uncertainty around estimated lagged correlations.
  
## Background
This repository contains an `R` package used by ZZZ (in preperation0) to qqq

## Installation
The current version of this package can be installed directly from github using
```{r,  eval = F}
devtools::install_github("ryanoisin/expct")
library(expct)
```

```{r,  eval = T, echo = F, message = F}
library(expct)
```

## Usage

The package takes as input a time-series dataset or long-format longitudinal data. The dataset should have the following columns
  - `id` a column denoting the id number of each participant in the dataset. Note that for single-subject time series, this should simply be a column with a single number repeated, e.g. `data$id = 1`
  - `time` a column containing timing information for each observation. The specific format required is "time elapsed since the **first** observation" in a unit of the users choice (hours, minutes, etc.). If the first observation of the dataset is taken at 3pm, and the second observation at 5pm, then the first two entries of the time column should read `time[1:2] = c(0,2)`. 
  - The remaining columns should contain records of the observed variables in long format.
  
```{r}
load("data/simdata.rda")
head(simdata)

```


The main function of this package is `expct`. Some important input options of this function are:

  - a) `dataset`: the dataset in the format described above
  - b) `Tpred`: A vector which indicates the time-intervals at which the user wants to estimate the auto- and cross- correlations. 
  - d) `output_type`: Determines the method used to construct credible or confidence intervals. If output_type == "CI", then default GAM confidence intervals are supplied, which can be interpreted as "point-wise" CIs for each of the values in `Tpred`. If output_type =="SCI", then simultaneous CIs are supplied. These can be interpreted as "function-wide" CIs, see [here](https://fromthebottomoftheheap.net/2016/12/15/simultaneous-interval-revisited/) for an introduction. If output_type =="LLCI", CIs are computed by approximating analytic auto- and cross-correlation standard errors from the time-series literature. The default value is "CI". We recommend using either "CI" or "SCI", with the latter being more conservative
  
The output of this functions is a list which contains below elements:

  - a) est: The point estimation of lagged correlations between covariates and outcomes.
  - b) highCI: The uppper CI of the point estimation.
  - c) lowCI: The lower CI of the point estimation.
  - d) laglongreducedummy: The stacked dataset from the raw data.
  - e) attributes: All attributes used to estimate GAMMs.  


### Example

Perform analyses with `expct`:

```{r, message = F}
# library(mgcv)
Tpred = seq(1,30,1)

est <- expct(
  dataset = simdata,
  Time = "time",
  outcome = c("Y1","Y2"),
  ID = "id",
  estimate = "marginal",
  Tpred = Tpred,
  plot_show = F,
  quantiles = c(.025, 0.975),
  method = "bam",
)
```

Plot output, for instance using
```{r, message = F}
plot(est$est[[1]], type = "b", ylab = "estimated autoregression", xlab = "time diff", main = "Autoregression X1")
lines(est$highCI[[1]], lty = 2, col = "gray")
lines(est$lowCI[[1]], lty = 2, col = "gray")

plot(est$est[[2]], type = "b", ylab = "estimated autoregression", xlab = "time diff", main = "Cross-Regression X2 on X1")
lines(est$highCI[[2]], lty = 2, col = "gray")
lines(est$lowCI[[2]], lty = 2, col = "gray")
# plot(mcr$est[[2]], type = "b", ylab = "estimated autoregression", xlab = "time diff", main = "Cross-Regresion")
# plot(mcr$est[[3]], type = "b", ylab = "estimated autoregression", xlab = "time diff", main = "Autoregression X1")
# plot(mcr$est[[4]], type = "b", ylab = "estimated autoregression", xlab = "time diff", main = "Autoregression X1")
```


## Contact Details

For more details please contact **o.ryan@uu.nl**
